\documentclass{beamer}

\input{settings.tex}


\title{Mechanics, Dynamics}
\subtitle{Mechatronics, Lecture 1}
\author{by Sergei Savin}
\centering
\date{\mydate}



\begin{document}
\maketitle



\begin{frame}{Content}
\begin{itemize}
\item Introduction
\item Newton - Euler laws
\item Moment of inertia
\item Mechanical Energy, Work, Power
\end{itemize}
\end{frame}


\begin{frame}{Introduction}
	%\framesubtitle{How do we know the state?}
	\begin{flushleft}
		
		
		What is Mechatronics? The term \emph{Mechatronics} is a combination of words \emph{mechanics} and \emph{electronics}.
		
		\bigskip
		
		\begin{definition}
			Mechatronics is a synergistic integration of mechanical and electrical engineering, computer control in design and manufacturing. 
		\end{definition}
		
		
	\end{flushleft}
\end{frame}


\begin{frame}{Introduction}
	%\framesubtitle{How do we know the state?}
	\begin{flushleft}
		
		
		Examples of mechatronic design include:
		
		\begin{itemize}
			\item Motors with built-in gears and sensors.
			
			\item Quadrotor brushless motors.
			
			\item Motor-wheel.
		\end{itemize}		
	
		\bigskip
		
		Modern robots, from robot-arm to Boston Dynamics's ATLAS are also examples of mechatronic design.
		
		
	\end{flushleft}
\end{frame}



\begin{frame}{Introduction}
	%\framesubtitle{How do we know the state?}
	\begin{flushleft}
		
		
		In this course we will focus on:
		
		\begin{itemize}
			\item Motors: working principles, control, sensing;
			
			\item Gears, reducers, transmissions: models, friction and other physical effects;
			
			\item Single input, single output (SISO) control, as applicable to simple electro-mechanical systems;
			
			\item Sensors, using sensor data.
			
		\end{itemize}		
		
		
	\end{flushleft}
\end{frame}



\begin{frame}{Newton - Euler law}
%\framesubtitle{How do we know the state?}
\begin{flushleft}

Second Newton's law in 2D case:

\begin{equation}
	\begin{cases}
		m \ddot x = \sum f_{i, x} \\
		m \ddot y = \sum f_{i, y}
	\end{cases}
\end{equation}
%
where $m$ is mass of a body, $f_{i, x}$, $f_{i, y}$ - Cartesian components of forces acting on the body. 

\end{flushleft}
\end{frame}



\begin{frame}{Newton - Euler law}
	%\framesubtitle{How do we know the state?}
	\begin{flushleft}
		
		Euler law of motion in 2D case:
		
		\begin{equation}
			J \ddot \varphi = \sum \tau_i
		\end{equation}
		%
		where $J$ is a moment of inertia of a body, $\varphi$ is orientation of the body, $\tau_i$ is torque. All of these entities are defined assuming rotation around the axis normal to the plane of motion, with positive direction of rotation defined as clock-wise.
		
		\bigskip
		
		We can see similarity of these equation with Newton laws. Note that this is only true for a 2D case, not for a general 3D case, where the Euler dynamics equations have the following form:
		
		\begin{equation}
			\bo{I} \dot{\omega} = \omega \times (\bo{I} \omega) + \tau
		\end{equation}
	%
	where $\bo{I}$ is matrix of inertia, $\omega$ is angular velocity and $\tau$ is external torque (all in body frame).
		
	\end{flushleft}
\end{frame}




\begin{frame}{Pendulum dynamics}
	%\framesubtitle{How do we know the state?}
	\begin{flushleft}
		
		A pendulum can be described by the following dynamical equations:
		
		\begin{equation}
			J \ddot \varphi = lmg \sin (\varphi)
		\end{equation}		
		%
		where $J$, $l$, $m$ and $g$ are moment of inertia, length, mass and gravitational acceleration. 
		
		\bigskip
		
		If there is motor torque $\tau$ acting on the pendulum, the dynamics takes form:
		
		\begin{equation}
			J \ddot \varphi = lmg \sin (\varphi) + \tau
		\end{equation}		
		
		
	\end{flushleft}
\end{frame}



\begin{frame}{Torque generated by a force}
	%\framesubtitle{How do we know the state?}
	\begin{flushleft}
		
		A force $\bo{f}$ applied to a rigid body at a point $\bo{r}$ generates a torque $\tau$:
		
		\begin{equation}
			\tau = \bo{r} \times \bo{f}
		\end{equation}	
		
		In 2D case, a force with magnitude $f$ acting along a line with a distance $r$ from the axis (around which torque is computed) generates torque according to the formula:
		
		\begin{equation}
			\tau = r f
		\end{equation}	
		
	\end{flushleft}
\end{frame}


\begin{frame}{Moment of inertia, 1}
%\framesubtitle{How do we know the state?}
\begin{flushleft}
	
	 It is tempting to think of moment of inertia in 2D case as an analog of mass; moment of inertia determines the rate of change of \textcolor{blue}{angular velocity} for a given \textcolor{blue}{torque}, same as mass determines the rate of change of \textcolor{mydarkturquoise}{linear velocity} for a given \textcolor{mydarkturquoise}{force}.
	 
	 \begin{equation*}
		\begin{cases}
			\textcolor{blue}{J \dot \omega =  \tau}  \\
			\textcolor{mydarkturquoise}{m \dot v = f}
	\end{cases}
\end{equation*}
	 
	
\end{flushleft}
\end{frame}




\begin{frame}{Moment of inertia, 2}
	%\framesubtitle{How do we know the state?}
	\begin{flushleft}
		
		Another way to think about moment of inertia is through kinetic energy. Kinetic energy of a rigid body can be described as:
		
		\begin{equation}
			T = \frac{1}{2} m \bo{v}\T \bo{v} + \frac{1}{2} \omega\T \bo{I} \omega
		\end{equation}			 
		
		In 2D case it is simplified:
		
		\begin{equation}
			T = \frac{1}{2} m (v_x^2 + v_y^2) + \frac{1}{2} J \omega^2
		\end{equation}		
		
		Moment of inertia determines how energy will the body acquire by the increase in its angular velocity.
		
	\end{flushleft}
\end{frame}



\begin{frame}{Moment of inertia, 3}
	%\framesubtitle{How do we know the state?}
	\begin{flushleft}
		
		Moment of inertia of a point mass $r$ meters away from the axis of rotation is given as:
		
		\begin{equation}
			J =m  r^2
		\end{equation}				
	
		The same formula give moment of inertia for a rim of a disk with radius $r$. For a solid disk with uniform mass distribution the formula becomes:
		
		\begin{equation}
			J =\frac{1}{2} m  r^2
		\end{equation}				
		
		For a centrally mounted rod the moment of inertia is $J =\frac{1}{12} m  r^2$ and for a rod mounted like a pendulum it is $J =\frac{1}{3} m  r^2$.
		
	\end{flushleft}
\end{frame}



\begin{frame}{Mechanical Energy}
	%\framesubtitle{How do we know the state?}
	\begin{flushleft}
		
		\emph{Mechanical energy} is a sum of kinetic and potential energy. We already saw the form of kinetic energy in 2D case: $T = \frac{1}{2} m (v_x^2 + v_y^2) + \frac{1}{2} J \omega^2$. 
		
		\bigskip
		
		Potential energy is more complex: it depends on what type of potential (conservative) forces that act on the system. For example:
		
		\begin{itemize}
			\item gravitational potential energy: $\Pi = mg (y - y_0)$, where $y_0$ is the value of vertical coordinate $y$ for which potential energy is defined as zero;
			
			\item linear spring potential energy: $\Pi = \frac{1}{2} k (r - r_0)^2$, where $k$ is the stiffness and $r_0$ is the length of the spring in the relaxed state.
		\end{itemize}
		
	\end{flushleft}
\end{frame}


\begin{frame}{Mechanical Work, 1}
	%\framesubtitle{How do we know the state?}
	\begin{flushleft}
		
		The change of the total energy of the system is equal to the sum of \emph{mechanical work} performed by the active forces:
		
		\begin{equation}
			\Delta T + \Delta \Pi = \sum A_i
		\end{equation}				
		%
		where $A_i$ is the work performed by the active force $f_i$ or a torque $\tau_i$.
		
		\bigskip
		
		We could compute the work performed by a force (applied to a point mass) on the interval of time from $t_0$ to $t_f$:
		
		\begin{equation}
			A = \int_{t_0}^{t_f} \bo{v}\T \bo{f} \  dt
		\end{equation}
		%
		where $\bo{v} = \bo{v}(t)$	is the velocity of the point mass and $\bo{f} = \bo{f}(t)$ is the force.
		
	\end{flushleft}
\end{frame}



\begin{frame}{Mechanical Work, 2}
	%\framesubtitle{How do we know the state?}
	\begin{flushleft}
		
		If a force $\bo{f}$ is collinear with velocity $\bo{v}$, we re-write the formula for mechanical work using the absolute values of these vectors $f$ and $v = \dot s$:
		%
		\begin{equation}
			A = \int_{t_0}^{t_f} v f \  dt
		\end{equation}
		
		As long as $f = \text{const}$, the expression becomes trivial:
		%
		\begin{equation}
			A =  (s(t_1) - s(t_0)) f
		\end{equation}
	
		We can derive the same formulas for mechanical work of a torque $\tau$:
		%
		\begin{align}
			A &= \int_{t_0}^{t_f} \omega \tau \  dt
			 \\
			A &= (\varphi(t_1) - \varphi(t_0)) \tau \ \ \ \textcolor{mygrey}{(\tau =  \text{const})} \
		\end{align}
	
	\end{flushleft}
\end{frame}


\begin{frame}{Mechanical Power}
	%\framesubtitle{How do we know the state?}
	\begin{flushleft}
		
		The entity $\bo{v}\T \bo{f}$ we saw before is called \emph{mechanical power}. It can also be defined as:
		
		\begin{align}
			P &= \bo{v}\T \bo{f}
			\\
			P &= \omega \tau
			\\
			P &= \frac{d}{dt} A
		\end{align}
		
		Note that where as mechanical work is an integral entity, mechanical power is instantaneous. Also, using the definition of mechanical energy we observe that:
		%
		\begin{equation}
			\frac{d}{dt} (T + \Pi) = \sum P_i
		\end{equation}	
		
	\end{flushleft}
\end{frame}


\myqrframe

\end{document}
